--- 
title: "DEW-landscapes"
link-citations: yes
bibliography:
-  "`r fs::path('/projects', 'data', 'bib', 'refs.bib')`"
-  "`r here::here('profile/packages.bib')`"
always_allow_html: yes
description: Background to DEW-landscapes packages and projects
output: github_document
---

```{r setup, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, tidy = TRUE, include = FALSE}

knitr::opts_chunk$set(echo = FALSE
                      , warning = FALSE
                      , error = FALSE
                      , message = FALSE
                      , tidy = TRUE
                      )

packages <- c("sits", "targets")

envFunc::check_packages(packages)

purrr::walk(packages, library, character.only = TRUE)

knitr::write_bib(file = here::here("profile/packages.bib"))

```

This is a work in progress. Last update `r format(Sys.time(), '%A, %d %B, %Y. %H:%M')`.

DEW-landscapes holds packages and projects we have found useful for creating, cleaning and analysing large, unstructured biological and environmental data sets in the context of providing lines of evidence for decision making in natural resources management.

This readme aims to:

* document the various `env` packages and projects
* specify the aims of each package and project
* point to useful supporting projects

Note that most projects are private github repositories. Links are provided to reports resulting from use of the projects, where available.

# Background

DEW-landscapes packages and projects have been roughly guided by the paper _Essential biodiversity variables for mapping and monitoring species populations_ [@RN4609] in that once data are accumulated, they are cleaned with reference to a regular cube of specific extent (generally with reference to temporal and geographic extent) and grain (generally with reference to temporal, geographic and taxonomic grain). Apparently, this aligns with the Earth Systems Data Cubes (ESDCs) concept [e.g. @RN4610]. The R package `sits` `r envReport::cite_package("sits", bib_file = here::here("profile", "packages.bib"))`, while not used directly in any `env` packages or projects, also provides useful background on [regular cubes](https://e-sensing.github.io/sitsbook/dc_regularize.html#the-need-for-regular-eo-data-cubes) within the context of earth observation (EO) cubes.

# Overview

The following figure shows a cartoon of the links between various `env` packages and projects.

```{r, results = "asis", echo = FALSE}

env_packages <- c("envImport", "envClean", "envRange", "envSDM")

mer <- targets::tar_mermaid(label = "description"
                            , outdated = FALSE
                            , color = FALSE
                            )

cat(c("```mermaid", mer, "```"), sep = "\n") 

if(FALSE) targets::tar_visnetwork()

```


# envImport and envOcc

The project [`envOcc`](https://github.com/dew-landscapes/envOcc) helps aggregate many disparate data sources. Functions for accessing publicly accessible data sources are available within [`envImport`](https://github.com/dew-landscapes/envImport).

# envClean and envCleaned


# Supporting packages

## [`envFunc`](https://github.com/dew-landscapes/envFunc)

`r utils::packageDescription("envFunc")$Description`

## [`envTargets`](https://github.com/dew-landscapes/envTargets)

`r utils::packageDescription("envTargets")$Description`

Specifically, in conjunction with supporting files `scales.yaml` and `_targets.yaml`, `envTargets` provides support for generating machine- and human-readable directory structures explicitly referencing the extent and grain 


# References


