--- 
title: "DEW-landscapes"
link-citations: yes
bibliography:
-  "`r fs::path('/projects', 'data', 'bib', 'refs.bib')`"
-  "`r here::here('profile/packages.bib')`"
always_allow_html: yes
description: Background to DEW-landscapes packages and projects
output: github_document
---

```{r setup, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, tidy = TRUE, include = FALSE}

knitr::opts_chunk$set(echo = FALSE
                      , warning = FALSE
                      , error = FALSE
                      , message = FALSE
                      , tidy = TRUE
                      )

packages <- c("sits", "targets")

envFunc::check_packages(packages)

purrr::walk(packages, library, character.only = TRUE)

knitr::write_bib(file = here::here("profile/packages.bib"))

```

This is a work in progress. Last update `r format(Sys.time(), '%A, %d %B, %Y. %H:%M')`.

DEW-landscapes holds packages and projects we have found useful for creating, cleaning and analysing large, unstructured biological and environmental data sets in the context of providing lines of evidence for decision making in natural resources management.

This readme aims to:

* document the various `env` packages and projects
* specify the aims of each package and project
* point to useful supporting projects

Package specific information is limited here - refer to the package repo for further description.

Note that most projects are private github repositories. Links are provided to reports resulting from use of the projects, where available.

# Background

DEW-landscapes packages and projects have been roughly guided by the paper _Essential biodiversity variables for mapping and monitoring species populations_ [@RN4609] in that once data are accumulated, they are cleaned with reference to a regular cube of specific extent (generally with reference to temporal and geographic extent) and grain (generally with reference to temporal, geographic and taxonomic grain). Apparently, this aligns with the Earth Systems Data Cubes (ESDCs) concept [e.g. @RN4610]. The R package `sits` `r envReport::cite_package("sits", bib_file = here::here("profile", "packages.bib"))`, while not used in any `env` packages or projects, also provides useful background on [regular cubes](https://e-sensing.github.io/sitsbook/dc_regularize.html#the-need-for-regular-eo-data-cubes) within the context of earth observation (EO) cubes.

# Overview

The following figure shows a cartoon of the links between various `env` packages and projects.

```{r, results = "asis", echo = FALSE}

mer <- targets::tar_mermaid(script = fs::path(here::here(), "_targets.R")
                            , label = "description"
                            , outdated = FALSE
                            , color = FALSE
                            , legend = FALSE
                            )

mer <- gsub("subgraph Graph", "subgraph Links between env packages and projects", mer)

cat(c("```mermaid", mer, "```"), sep = "\n") 

if(FALSE) targets::tar_visnetwork()

```

# Assemling

## [`envImport`](https://github.com/dew-landscapes/envImport)

A package with description: `r utils::packageDescription("envImport")$Description`

## [`envOcc`](https://github.com/dew-landscapes/envOcc)

The project `envOcc` aggregates occurrence data from many disparate data sources.

## [`envVec`](https://github.com/dew-landscapes/envVec)

This project generates spatial vector data (i.e. point, line and polygon layers, as geoparquet files) for use in other env projects.

## [`envRaster`](https://github.com/dew-landscapes/envRaster)

A package with description: `r utils::packageDescription("envRaster")$Description`

## [`envRas`](https://github.com/dew-landscapes/envRas)

This project generates environmental data as rasters, for use in combination with the biological data assembled via `envOcc`.

## [`envFire`](https://github.com/dew-landscapes/envFire)

A package with description: `r utils::packageDescription("envFire")$Description`

# Cleaning

## [`envClean`](https://github.com/dew-landscapes/envClean)

A package with description: `r utils::packageDescription("envClean")$Description`

`envClean` functions are most obviously applied in the `envCleaned` workflow. However, many `envClean` functions will be applied only in downstream projects as they are only useful with respect to a specific projects aims. For example, records taken of only a single taxa within a location (say, as part of tree health monitoring) provide useful data for a species distribution model but do not provide useful data for community analyses. Thus the function [`envClean::make_effort_mod()`](https://dew-landscapes.github.io/envClean/reference/make_effort_mod.html) &ndash; that allows detecting excessively over- or under-sampled visits for possible filtering &ndash; is unlikely to be useful in preparing species distribution models but is essential when undertaking community analyses.

## [`envDistribution`](https://github.com/dew-landscapes/envDistribution)

A package with description: `r utils::packageDescription("envDistribution")$Description`

## [`envCleaned`](https://github.com/dew-landscapes/envCleaned)

The project `envCleaned` takes the data assembled by `envOcc` and applies generic cleaning steps that will be relevant to any downstream use.

Some example generic cleaning steps include:

* [`envClean::make_taxonomy()`](https://dew-landscapes.github.io/envClean/reference/make_taxonomy.html). Query online databases to retrieve a good guess (at best!) for standardised taxonomy for use across all the assembled data sources.
* [`envClean::filter_geo_range()`](https://dew-landscapes.github.io/envClean/reference/filter_geo_range.html). Can be called directly on a dataframe to filter records to those occurring within a polygon (or its bounding box).
* [`envDistribution::filter_by_distribution()`](https://dew-landscapes.github.io/envDistribution/reference/filter_by_distribution.html). Can be used to flag or filter records of a taxa outside of its range, as specified by a polygon.

## [`envStatus`](https://github.com/dew-landscapes/envStatus)

Generates taxonomically aligned (using 'envClean::make_taxonomy') information on taxa listed under various legislation.

# Main outputs

## [`envRegCont`](https://github.com/dew-landscapes/envRegCont)

What contribution does any region (represented by a polygon) make to the distribution (usually Australian) of a taxa (based on IUCN Red List AOO and EOO metrics)?

## [`envSDMs`](https://github.com/dew-landscapes/envSDMs)

Generate many, many species distribution models (SDMs). Generates outputs that sit somewhere between dots-on-a-map and broad, envelope-type distribution models. In many cases, the models show sub-landscape preference. For example, _Sminthopsis psammophila_ more likely to be found on sand dunes in landscapes of sandy dunes and calcareous-loam swales.

## [`envPIA`](https://github.com/dew-landscapes/envPIA)

For any area of interest (aoi) generate maps of areas that are more or less likely to trigger various biodiversity legislation via the following process:

1) Generate an initial inclusive list of binomials for any aoi (aoi-taxa), made up of any taxa with:
    1) records within the aoi, including a buffer
    1) one of the following overlaps the study area
        1) an existing _Environment Protection and Biodiversity Conservation Act 1999 (EPBC Act)_ distribution that overlaps the study area, not including a buffer, or, if not available
        1) an existing expert distribution generated for some fauna (about 70 fauna taxa), including a buffer, or, if not available
        1) a minimum convex polygon (mcp) around all suitable records of a taxa, including a buffer
1) Identify taxa of concern (aoi-concern), made up of any aoi-taxa:
    1) listed as threatened by the _EPBC Act_
    1) listed on schedules 7 (Endangered), 8 (Vulnerable) or 9 (Rare) of the _National Parks and Wildlife Act 1972 (NPW Act)_
    1) likely to be listed in the future under the _EPBC Act_ or _NPW Act_
    1) with a significant proportion of their Australian distribution in the aoi (see [`envRegCont`](https://github.com/dew-landscapes/envRegCont/))
1) Provide a fine-scale representation of where, within the aoi, aoi-concern taxa are likely to occur based on (in order):
    1) fine-scale distribution model for each taxa (species distribution model - sdm)
    1) if no fine_scale distribution model could be developed, then based on EPBC Act distribution
    1) if no fine-scale distribution model could be developed, and the taxa had no EPBC Act distribution, then based on association between records and an appropriate environmental layer
1) Summarise the fine-scale representations into counts of taxa:
    1) listed on the _EPBC Act_
    1) listed on schedules 7, 8 or 9 of the _NPW Act_
    1) likely to be listed in future
    1) that are not listed on any legislation but that have a significant proportion of their Australian distribution in the aoi

## [`envEco`](https://github.com/dew-landscapes/envEco)

For any area of interest (aoi) generate maps and descriptions of ecosystems.

## [`envSens`](https://github.com/dew-landscapes/envSens)

Based on species trait data, estimate sensitivity to stresses.

## `sa_va`

Build trait database for fire-relevant traits for South Australian taxa.

## `fire_sensitivity`

Preferred fire return interval for South Australian ecosystems (based on `envEco`) using trait information from `sa_va`.

## [`envDetection`](https://github.com/dew-landscapes/envDetection)

Generates detection estimates associated with different sampling methods for taxa via occupancy models. Used to determine the relevant sampling methods for each taxa to improve target group sampling, and generation of bias layers and background points in SDMs. 

## [`DFaB`](https://github.com/Calamanthus/DFaB)

Dynamic Fire and Biodiversity tool for managing habitat of fire dependent fauna via burning. 

## `envRegAssess`

IUCN Red List regional assessments for a region based on Criterion B (i.e. regional AOO & EOO estimates and associated threat ratings).  

# Supporting packages

## [`envFunc`](https://github.com/dew-landscapes/envFunc)

`r utils::packageDescription("envFunc")$Description`

## [`envTargets`](https://github.com/dew-landscapes/envTargets)

`r utils::packageDescription("envTargets")$Description`

Specifically, in conjunction with supporting files `scales.yaml` and `_targets.yaml`, `envTargets` provides support for generating machine- and human-readable directory structures explicitly referencing the extent and grain

## [`envShiny`](https://github.com/dew-landscapes/envShiny)

`r utils::packageDescription("envShiny")$Description`

Again with explicit support around extent and grain.


# References

